# Regression

Use example:

```{r message=FALSE}
library(tidyverse)
reg_data <- read_csv("https://raw.githubusercontent.com/my1396/course_dataset/refs/heads/main/META_monthly_factor_model_2014-2024.csv")
reg_data <- reg_data %>% 
    mutate(return = adjusted/lag(adjusted)-1,
           eRi = return-RF)
## CAPM
capm_ml <- lm(eRi~rmrf, data=reg_data)
summary(capm_ml)
```

Get the **coefficient table** with `broom::tidy`.

```{r}
library(broom)
tidy(capm_ml)
```


Get the **variance-covariance matrix** with `vcov`.

```{r}
vcov(capm_ml)
# verify using coef table
capm_ml %>% 
    tidy() %>% 
    mutate(variance=std.error^2)
```


We can manually calculate as
```{r}
df <- (nrow(capm_ml$model)-2)  # degree of freedom
sigma2 <- sum(capm_ml$residuals^2)/df  # residual variance
# 1st column of lm$model is the depend. var.
X <- capm_ml$model[, 2, drop=FALSE] %>% as.matrix()
X <- cbind(1, X)  # add intercept
V <- solve(t(X) %*% X) * sigma2
V
```


Note that `cov.unscaled` returns the **unscaled** covariance matrix, $(X'X)^{-1}$. 

- To get estimated covariance matrix for the coefficients, you need to multiply `cov.unscaled` by the estimate of the error variance.

```{r}
solve(t(X) %*% X)
summary(capm_ml)$cov.unscaled
all.equal(solve(t(X) %*% X), 
          summary(capm_ml)$cov.unscaled, 
          check.attributes = FALSE)
```


**References**:

- <https://github.com/SurajGupta/r-source/blob/master/src/library/stats/R/vcov.R>
- stats package: <https://docs.tibco.com/pub/enterprise-runtime-for-R/6.0.1/doc/html/Language_Reference/stats/summary.lm.html>






